### 2015-03-11 src/3_TimeSeries.tex
adding AVECM proposal

The vector autoregressive VAR($p$) model is one of the most easy to use, successful and flexible models for the analysis of multivariate time series. It is a natural extension of the univariate autoregressive (AR) model. The VAR model has proven to be useful for describing the dynamic behaviour of economic and financial time series.
VAR is a general framework describing the behaviour of a
set of $l$ endogenous variables as a linear combination of their last $p$
values, where $l,p\in\mathbb{N}$. 
In our case, each one of these $l$ variables is a scalar time series
$y_{\lambda,t}$, $\lambda=1,\dots,l$, and we represent them all together
at time $t$ by the the vector time series:
\begin{equation}
\label{eq:variables}
\mathbf{y}_t = 
\begin{bmatrix} y_{1,t} & y_{2,t} & \dots & y_{l,t} \end{bmatrix}^\top.
\end{equation}
\noindent
Notice that the vectors $\mathbf{y}_t$ are assumed to be $l$-dimensional.

@@

\section{Vector error correction model}
Vector error correction model (VECM) is a special form of a VAR model for I(1) variables that are also
cointegrated~\cite{banerjee1993}.

@@

VECM models are employed because many economic time series appear to be first-difference stationary, i.e. I(1)

@@

If two series are both integrated (of order one, or I(1)) we could model
their interrelationship by taking first differences of each series and
including the differences in a VAR or a structural model.
However, this approach would be suboptimal if it was determined that
these series are indeed cointegrated. In that case, the VAR only
express the short-run responses of these series to innovations in each
series. This implies that the simple regression in first differences is
misspecified.
If the series are cointegrated, they move together in the long-run. A
VAR in first differences will not capture those long-run tendences.
VECM fit to the first differences of the non stationary variables, but a lagged error-correction term is added to the relationship representing the long-run relationship.

>>>>

The vector autoregressive VAR($p$) model \cite{sims1980} is one of the most easy to use, successful and flexible models for the analysis of multivariate time series. It is a natural extension of the univariate autoregressive (AR) model. The VAR model has proven to be useful for describing the dynamic behaviour of economic and financial time series.
VAR is a general framework describing the behaviour of a
set of $l$ endogenous variables as a linear combination of their last $p$
values, where $l,p\in\mathbb{N}$. 
In our case, each one of these $l$ variables is a scalar time series
$y_{\lambda,t}$, $\lambda=1,\dots,l$, and we represent them all together
at time $t$ by the the vector time series:
\begin{equation}
\label{eq:variables}
\mathbf{y}_t = 
\begin{bmatrix} y_{1,t} & y_{2,t} & \dots & y_{l,t} \end{bmatrix}^\top.
\end{equation}
\noindent
Notice that the vectors $\mathbf{y}_t$ are assumed to be $l$-dimensional.

@@

\section{Vector error correction model}
Vector error correction model (VECM)  \cite{engle87} is a special form of a VAR model for I(1) variables that are also
cointegrated~\cite{banerjee1993}.

@@

@@

If two series are both integrated (of order one, or I(1)) we could model
their interrelationship by taking first differences of each series and
including the differences in a VAR or a structural model.
However, this approach would be suboptimal if it was determined that
these series are indeed cointegrated. In that case, the VAR only
express the short-run responses of these series to innovations in each
series. This implies that the simple regression in first differences is
misspecified.
If the series are cointegrated, they move together in the long-run. A
VAR in first differences will not capture those long-run tendencies.
VECM fit to the first differences of the non stationary variables, but a lagged error-correction term is added to the relationship representing the long-run relationship.

In finance, many economic time series are revealed to be stationary when they
are differentiated and moreover cointegration restrictions often improve
forecasting \cite{duy1998}. Therefore, VECM has been widely adopted in
finance: \cite{mukherjee1995}, \cite{maysami2000}, \cite{arestis2001} and
\cite{seong2013} to name a few. Pair
trading is a very common example of a cointegration
application \cite{herlemont2003} but also can also be extended to a larger set
of variables \cite{mukherjee1995}, \cite{engle2004}.
